Analytic Occultation Light Curves in Reflected Light

### 2020-04-03 tex/ms.tex
more text, but frame notation is messed up

%
\section{The Problem}
\label{sec:the-problem}
%
\subsection{Review of the \starry algorithm in emitted light}
\label{sec:starry-review}
%
Without loss of generality, assume the body whose flux we wish to compute
has radius unity and sits at the origin of a right-handed Cartesian coordinate
system with the $x$-axis pointing to the right on the sky, the $y$-axis
pointing up on the sky, and the $z$-axis pointing out of the sky
toward the observer; denote this frame by $\mathcal{F}_0$. In $\mathcal{F}_0$,
the surface (emitted) intensity field of the body is described by the
spherical harmonic coefficient vector $\mathbf{y}$. Following
\citet{Luger2019}, when no occultor is present, we may compute the total
visible flux $f_0$ from this body as
%
\begin{align}
    \label{eq:rTA1Ry}
    f_0 = \rTe \BF{A_1} \BF{R} \BF{y}
    \quad,
\end{align}
%
where, from right to left, $\BF{R} = \BF{R}(i, \lambda, \vartheta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ to the sky frame
$\mathcal{F}_\mathrm{sky}$ given the body's inclination $i$, obliquity
$\lambda$, and rotational phase $\vartheta$ (Equation~\ref{eq:R}),
%
$\BF{A_1}$ is the change-of-basis matrix from the spherical harmonic
basis $\by$ to the \emph{polynomial basis} $\bp$ in which the integrals
are computed (Equation~\ref{eq:A1}),
%
and $\rTe$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bp$
(Equation~\ref{eq:rTe}).
%
If instead an occultor of radius $r_o$ is located at position $(x_o, y_o)$,
we instead compute the visible flux $f$ from
%
\begin{align}
    \label{eq:sTARRy}
    f = \sTe \BF{A} \BF{R}' \BF{R} \BF{y}
    \quad,
\end{align}
%
where, as before, $\BF{R}$ rotates the body from $\mathcal{F}_0$
to $\mathcal{F}_\mathrm{sky}$,
%
$\BF{R}' = \BF{R}'(\omega)$ rotates the body on the plane
of the sky by an angle
$\omega = \nicefrac{\pi}{2} - \mathrm{arctan2}(y_o, x_o)$
into a frame $\mathcal{F}_\mathrm{greens}$ in which the
occultor lies along the $+y$-axis (Equation~\ref{eq:R'}),
%
$\BF{A}$ is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed
(Equation~\ref{eq:A}),
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
(Equation~\ref{eq:sTe}), with $b_o = \sqrt{x_o^2 + y_o^2}$.
For more details, see Appendix~\ref{app:starry} and \citet{Luger2019}.

@@

$+x$-axis,

@@

    \theta = \mathrm{arctan2}(y_s, x_s)

@@

Recalling the structure of the basis (Equation~\ref{eq:p}),
    \bp & =
        1   &
        x   & z  & y  &
        x^2 & xz & xy & yz & y^2 &
        \cdot\cdot\cdot
    \end{pmatrix}^\mathsf{T}

@@

Figure~\ref{fig:frames} summarizes the transformations involved in the two
equations above. Starting on the right with a map vector $\BF{y}$ in
the spherical harmonic basis $\by$, defined in some observer-independent frame
$\mathcal{F}_0$, we first rotate it via $\BF{R}$ to the sky frame
$\mathcal{F}_\mathrm{sky}$, in which orientation the body is viewed by the
observer. If there is no occultation (upper branch of the figure), we then
rotate the map once more via $\BF{R}'$ to the integration frame
$\mathcal{F}_\mathrm{int}$, in which the terminator is parallel to the
$x$-axis. We then apply $\BF{A_1}$ to change basis to $\bp$, apply the
illumination transform $\BF{I}$, and finally dot in the solutions to the
surface integrals $\rT$. If, on the other hand, an occultor is present,
we rotate the map from $\mathcal{F}_\mathrm{sky}$ via $\BF{R}'$ to the frame
$\mathcal{F}_\mathrm{greens}$, in which the occultor lies along the
$+y$-axis. We then apply $\BF{A_1}$ to change basis to $\bp$ and $\BF{I}$
to weight the map by the illumination as above. Finally, we change basis
via $\BF{A_2}$ to the Green's basis, in which we compute and dot the
integrals $\sT$.

@@

$\mathcal{F}_\mathrm{int}$

@@

recalling that in frame $\mathcal{F}_\mathrm{int}$, $\theta = 0$ by

@@

Figure~\ref{fig:cases} shows the
eight families of cases involving an occultation of an illuminated body by
a spherical occultor. Each of these cases correspond to distinct ways in
which the limb of the occultor can intersect with the limb of the occulted
body and its terminator; together, these cases encompass all possible
occultation configurations, for any illumination angle, occultor size, and
occultor position. In the following sections, we present the solution for the
$\sT$, and hence for the total visible flux, for each of these cases.

\xxx{How do we determine which case we're modeling?}

@@

\section{The \starry formalism}
\label{app:starry}
%
The basis $\bg$ is called the \emph{Green's basis}; its name
stems from the fact that its components have a structure that makes integration
by Green's theorem convenient. Its components were defined in
Equation~11 of \citet{Luger2019b}:
%
\begin{proof}{bg}
    \tilde{g}_{l,m} &=
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \mathrm{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \, \mathrm{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \mathrm{otherwise}
        \quad,
    \end{dcases}
    \label{eq:bg}
\end{proof}
%
where $x$ and $y$ are the Cartesian coordinates on the surface
of the projected
disk of the body, with $z(x, y) = \sqrt{1 - x^2 - y^2}$, and
%
\begin{align}
    \label{eq:munu}
    \mu & = l - m
    \nonumber     \\
    \nu & = l + m
    \quad.
\end{align}
%
The vector $\bg$ is ordered such that
the component at index $n$ is $\tilde{g}_{l,m}$, with
%
\begin{align}
    \label{eq:lm}
    l & = \floor*{\sqrt{n}} \nonumber \\
    m & = n - l^2 - l
    \quad.
\end{align}
%
Finally, the rotation matrix $\BF{R} = \BF{R}(i, \lambda, \vartheta)$
rotates the body to the correct orientation on the sky given its
inclination $i$, obliquity $\lambda$, and rotational phase $\vartheta$,
while $\BF{R}'$ rotates the body on the plane
of the sky into the frame $\mathcal{F}'$ in which the integration is
actually performed, with the occultor along the $+y$-axis.
For more information on all of these terms, the reader is referred to
\citet{Luger2019}.

>>>>

%
\section{The Problem}
\label{sec:the-problem}
%
\textbf{Note:}
This section closely follows the
notation and formalism introduced in \citet{Luger2019}. While we
include all of the relevant equations and definitions in
Appendix~\ref{app:starry}, the reader is strongly encouraged to
review \S2--3 of \citet{Luger2019} before proceeding.

\subsection{Review of the \starry algorithm in emitted light}
\label{sec:starry-review}
%
Without loss of generality, assume the body whose flux we wish to compute
has radius unity and sits at the origin of a right-handed Cartesian coordinate
system with the $x$-axis pointing to the right on the sky, the $y$-axis
pointing up on the sky, and the $z$-axis pointing out of the sky
toward the observer; denote this frame by $\mathcal{F}_0$. In $\mathcal{F}_0$,
the surface (emitted) intensity field of the body is described by the
spherical harmonic coefficient vector $\mathbf{y}$. Following
\citet{Luger2019}, when no occultor is present, we may compute the total
visible flux $f_0$ from this body as
%
\begin{align}
    \label{eq:rTA1Ry}
    f_0 = \rTe \BF{A_1} \BF{R} \BF{y}
    \quad,
\end{align}
%
where, from right to left, $\BF{R} = \BF{R}(i, \lambda, \vartheta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ to the sky frame
$\mathcal{F}_\mathrm{sky}$ given the body's inclination $i$, obliquity
$\lambda$, and rotational phase $\vartheta$ (Equation~\ref{eq:R}),
%
$\BF{A_1}$ is the change-of-basis matrix from the spherical harmonic
basis $\by$ to the \emph{polynomial basis} $\bp$ in which the integrals
are computed (Equations~\ref{eq:by}, \ref{eq:bp}, and \ref{eq:A1}),
%
and $\rTe$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bp$
(Equation~\ref{eq:rTe}).
%
If instead an occultor of radius $r_o$ is located at position $(x_o, y_o)$,
we instead compute the visible flux $f$ from
%
\begin{align}
    \label{eq:sTARRy}
    f = \sTe \BF{A} \BF{R}' \BF{R} \BF{y}
    \quad,
\end{align}
%
where, as before, $\BF{R}$ rotates the body from $\mathcal{F}_0$
to $\mathcal{F}_\mathrm{sky}$,
%
$\BF{R}' = \BF{R}'(\omega)$ rotates the body on the plane
of the sky by an angle
$\omega = \mathrm{arctan2}(x_o, y_o)$
into a frame $\mathcal{F}_\mathrm{greens}$ in which the
occultor lies along the $+y$-axis (Equation~\ref{eq:R'}),
%
$\BF{A}$ is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed
(Equation~\ref{eq:A}),
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
(Equation~\ref{eq:sTe}), with $b_o = \sqrt{x_o^2 + y_o^2}$.
For more details, see Appendix~\ref{app:starry} and \citet{Luger2019}.

@@

$+x$-axis \xxx{CHECK ME},

@@

    \theta &=
        -\mathrm{arctan2}(x_s, y_s)
         & \qquad \mathcal{F} = \mathcal{F}_\mathrm{term}   \\
        \omega - \mathrm{arctan2}(x_s, y_s)
         & \qquad \mathcal{F} = \mathcal{F}_\mathrm{greens}

@@

Recalling the structure of the basis (Equation~\ref{eq:bp}),

@@

Figure~\ref{fig:frames} summarizes the transformations involved in the two
equations above. Starting on the right with a map vector $\BF{y}$ in
the spherical harmonic basis $\by$, defined in some observer-independent frame
$\mathcal{F}_0$, we first rotate it via $\BF{R}$ to the sky frame
$\mathcal{F}_\mathrm{sky}$, in which orientation the body is viewed by the
observer. If there is no occultation (upper branch of the figure), we then
rotate the map once more via $\BF{R}'$ to the integration frame
$\mathcal{F}_\mathrm{term}$, in which the terminator is parallel to the
$x$-axis. We then apply $\BF{A_1}$ to change basis to $\bp$, apply the
illumination transform $\BF{I}$, and finally dot in the solutions to the
surface integrals $\rT$. If, on the other hand, an occultor is present,
we rotate the map from $\mathcal{F}_\mathrm{sky}$ via $\BF{R}'$ to the frame
$\mathcal{F}_\mathrm{greens}$, in which the occultor lies along the
$+y$-axis. We then apply $\BF{A_1}$ to change basis to $\bp$ and $\BF{I}$
to weight the map by the illumination as above. Finally, we change basis
via $\BF{A_2}$ to the Green's basis, in which we compute and dot the
integrals $\sT$.

@@

$\mathcal{F}_\mathrm{term}$

@@

recalling that in frame $\mathcal{F}_\mathrm{term}$, $\theta = 0$ by

@@

The integration in the unocculted case presented above is relatively
straightforward, since
the boundaries of integration are always the half-ellipse defining the
terminator and the half-circle defining the upper limb of the body
(Equation~\ref{eq:rT}). When an occultor is present, however, the
integration boundaries are far less trivial, since they may or may not
include sections of the terminator, sections of the limb of the body,
and sections of the limb of the occultor. There are in total eight families of
configurations, each defined by a distinct combination of integration
boundaries; these are shown in Figure~\ref{fig:cases}.
Together, these cases encompass all possible
occultation configurations, for any illumination angle, occultor size, and
occultor position.

Before we discuss how to compute the occultation integrals, we must first
develop a procedure to identify the relevant case given the occultor
position $(x_o, y_o)$ and radius $r_o$ and the terminator semi-minor
axis $b$ and angle $\theta$. Then, once the case is determined, we must
identify the relevant integration boundaries, which depend on the points
of intesection between the limb of the body, the limb of the occultor, and
the terminator. We do so in the following section.

\subsection{Case determination and integration boundaries}
\label{sec:which-case}
%
The key to identifying the case corresponding to a given configuration is
to determine whether or not the limb of the occultor intersects the terminator
of the body, and if so, the points of intersection.
Switching momentarily to the frame $\mathcal{F}_\mathrm{term}$,
the equations defining the
\begin{align}
    y_1(x) & = b \sqrt{1 - x^2}
    \nonumber                                       \\
    y_2(x) & = y'_o \pm \sqrt{r_o^2 - (x - x'_o)^2}
\end{align}
%
where $x'_o = $ and $y'_o = $ are the coordinates of the occultor in
$\mathcal{F}_\mathrm{term}$.

Recall that we perform the integration in frame
$\mathcal{F}_\mathrm{greens}$, in which the occultor lies along the $+y$-axis
at $(0, b_o)$, with $b_o = \sqrt{x_o^2 + y_o^2}$. In this frame, the
semi-major axis of the terminator is at an angle $\theta$ with respect to the
$+x$-axis (Equation~\ref{eq:theta}).

@@

\section{The \starry formalism}
\label{app:starry}
%
Equations:
%
\begin{align}
    \label{eq:R}
    \BF{R} = ?
\end{align}
%
\begin{align}
    \label{eq:by}
    \by = ?
\end{align}
%
\begin{align}
    \label{eq:bp}
    \bp & =
    \begin{pmatrix}
        1   &
        x   & z  & y  &
        x^2 & xz & xy & yz & y^2 &
        \cdot\cdot\cdot
    \end{pmatrix}^\mathsf{T}
    \quad,
\end{align}
%
\begin{align}
    \label{eq:A1}
    \BF{A_1} = ?
\end{align}
%
\begin{align}
    \label{eq:rTe}
    \rTe = ?
\end{align}
%
\begin{align}
    \label{eq:R'}
    \BF{R}' = ?
\end{align}
%
\begin{align}
    \label{eq:A}
    \BF{A} = ?
\end{align}
%
\begin{align}
    \label{eq:sTe}
    \sTe = ?
\end{align}
%
The basis $\bg$ is called the \emph{Green's basis}; its name
stems from the fact that its components have a structure that makes integration
by Green's theorem convenient. Its components were defined in
Equation~11 of \citet{Luger2019b}:
%
\begin{proof}{bg}
    \tilde{g}_{l,m} &=
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \mathrm{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \mathrm{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \mathrm{odd}, \,
        \mu = 1, \,
        \, \mathrm{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \mathrm{otherwise}
        \quad,
    \end{dcases}
    \label{eq:bg}
\end{proof}
%
where $x$ and $y$ are the Cartesian coordinates on the surface
of the projected
disk of the body, with $z(x, y) = \sqrt{1 - x^2 - y^2}$, and
%
\begin{align}
    \label{eq:munu}
    \mu & = l - m
    \nonumber     \\
    \nu & = l + m
    \quad.
\end{align}
%
The vector $\bg$ is ordered such that
the component at index $n$ is $\tilde{g}_{l,m}$, with
%
\begin{align}
    \label{eq:lm}
    l & = \floor*{\sqrt{n}} \nonumber \\
    m & = n - l^2 - l
    \quad.
\end{align}
%
Finally, the rotation matrix $\BF{R} = \BF{R}(i, \lambda, \vartheta)$
rotates the body to the correct orientation on the sky given its
inclination $i$, obliquity $\lambda$, and rotational phase $\vartheta$,
while $\BF{R}'$ rotates the body on the plane
of the sky into the frame $\mathcal{F}'$ in which the integration is
actually performed, with the occultor along the $+y$-axis.
For more information on all of these terms, the reader is referred to
\citet{Luger2019}.
